---
title: "Notes"
author: "Hursh Gupta"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Some Basics

-   Use `Ctrl` + `Enter` to run a code line under the cursor
-   R doesn't consider **next line**, it uses *strict parenthesis*
-   Some useful R commands

```{r packages, include=FALSE}
#q()
#help.start()
#help(func_name)
#example(func_name)
#help.search(item)
#help(package = "ggplot2")
## vignettes are additional documents such as tutorials, intros etc.
#vignette()
#vignette(package = "ggplot2")
```


### Everything in **R** is basically an object. There are *6 basic* data types

-   character "a" "z"
-   numeric (real or decimal) $3,12.2$ (default)
-   integer $5L$ (L is an instruction to store as an int)
-   logical TRUE FALSE
-   complex $1 + 2i$

## Class Vs TYPE vs MODE

```{r meta}
#class()
#typeof()
#length()
#attributes()
nchar("sairam hi there")
```

## Printing

```{r printing}
pi = sqrt(5)
print(pi)
```

-   print takes care of the formatting
-   can print only one obj at a time

e.g.

```{r catenation}
#print("the value of pi is ", pi,"in two decimals")
# the above doesn't work, use what's below

print(a <- cat("the value of pi is ", pi,"in two decimals"))
```

## Assignment

```{r variables}
x <- 3
y <- 5
print(sqrt(x^2 + y^3))
```

-   variable held in workspace; we can change variable's data type at will

```{r more_variables}
x <- "sunday"
print(x)
x <<- 5
# global variable
```

`=` assignment; `->` can also be used *(check the difference)*

### CHECK the scope

```{r scope_variables}
ls()
ls.str()
# the str() is called for showing the variable structure
```

```{r}
rm(x)
rm (list = ls()) #dangerous
```

# Data Structures

Understand as atomic vectors and **not lists**

## Types:

-   character
-   logical
-   integer
-   numeric
-   empty vector

```{r vectors}
a <- vector();a
b <- vector("logical", length = 3);b

logical(5)
```

```{r c_function}
vec1 = c(1,2,5)
# c is short for combine

typeof(vec1)
vec2 = c("sun","mon")
vec3 = c(vec1,vec2)

```

A Vector should only have **similar type elements**.

```{r type_casting}
vec4 = c(vec2, as.character("3")); vec4
a <- c(1,0,as.logical(3));a
a <- c(1,0, as.numeric(FALSE));a
a <- c(TRUE, FALSE, as.numeric(2));a
```

Again, a vector cannot contain mix data types.\\ Data elements can co-exist in a vector only if they have the some mode.

```{r NA}
x <- c("mon",NA, "wed")
```

## Comparison Operators


When you compare a multi-element array to a single element it compares all the elements to that single element and returns a logical vector. 

```{r comparison}
# comparing  == != < > <= >=
c(1,2,3,pi)==pi
```